---
title: "Text Analysis"
editor: visual
---
I got this data from the Kaggle website https://www.kaggle.com/datasets/fabriziocominetti/the-office-lines
The direct link is not longer active, https://www.officequotes.net/ 
However, the cited source to the TV show is The Office. Created by Greg Daniels, performances by Steve Carell, Rainn Wilson, and John Krasinski, NBC, 2005–2013.
```{r, include=FALSE}
library(tidyverse)
```

```{r}
office <- read.csv("the-office_lines.csv")
```



```{r}

allline<-office %>% 
  mutate(Line2 = str_split(Line, " ")) %>% 
  unnest_longer(Line2) %>%
  mutate(Line2 = str_to_lower(Line2))
library(stringr)
pat <- regex("\\bjesus\\b\\s+\\bchrist\\b", ignore_case = TRUE)
total <- sum(str_count(office$Line, pat), na.rm = TRUE)
total

```

```{r}
allline<-office %>% 
  mutate(Line2 = str_split(Line, " ")) %>% 
  unnest_longer(Line2) %>%
  mutate(Line2 = str_to_lower(Line2))
library(stringr)
allline$Line2 <- as.character(allline$Line2)
pat <- regex("\\bjesus\\b(?!\\s+christ\\b)", ignore_case = TRUE)
counts <- str_count(allline$Line2, pat)
total_not_jesus_christ <- sum(counts, na.rm = TRUE)
total_not_jesus_christ
```


```{r}
library(ggplot2)

df <- data.frame(
  group = c("Jesus Christ", "Jesus"),
  value = c(2, 29)
)

ggplot(df, aes(x = group, y = value)) +
  geom_point(shape = 3, size = 6, stroke = 1.5, color = "gold") +
  labs(title = "How Many Times Jesus Was Said In The Office", 
       x = "Phrase Said", y = "Number of Times Said") 


```

This graph illustrates how many times the word "Jesus" was said in "The Office." The insights I gained from my plot is how often "Christ" follows "Jesus." The graph shows the frequency of the collocation. It is suprising how uncommon it occurs.

```{r}
plot2<-office %>% mutate(Line2 = str_split(Line, " ")) %>% 
  unnest_longer(Line2)
```

```{r}
plot3 <- plot2 %>% mutate(office = str_count(Line2, regex("\\boffice\\b"))) %>% group_by(Character) %>% 
  summarise(word = n(), office = sum(office)) %>% filter(office > 0)
head(plot3)
```

```{r}
library(tidyverse)
main_chars <- c(
  "Michael", "Dwight", "Jim", "Pam", "Andy", "Ryan", "Stanley",
  "Phyllis", "Angela", "Kevin", "Oscar", "Meredith", "Creed",
  "Toby", "Kelly", "Erin", "Darryl"
)
plot2_main <- plot3 %>%
  filter(Character %in% main_chars) %>%
  mutate(label = paste0(Character, " (", office, ")"))  
plot2_main %>%
  ggplot(aes(x = "", y = office, fill = label)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c(
    "#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
    "#FF7F00", "#FFFF33", "#A65628", "#F781BF",
    "#999999", "#66C2A5", "#FC8D62", "#8DA0CB",
    "#E78AC3", "#A6D854", "#FFD92F", "#E5C494", "#00FF66"
  )) +
  labs(title = "Times Each Main Office Character Said 'Office'") +
  theme_void() +
  theme(
    legend.title = element_blank(),
  )

```

This graph illustrates how many times the main characters of the "The Office" said office. Besides the pie chart, there is also a table with character names and the frequency of how often they said the word. The insights I gained from my plot is Michael, the main character, is the most frequent user of the word "office", but no character makes up the majority.
